Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks

### 2016-11-04 docs/papers/2016JNCA/2016JNCA.diff1.tex
first round of corrections

\section{Complexity Analysis}
\label{sec:Complexity}

This section discusses the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.

@@

@@

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack.

@@

falar sobre as melhorias para reduzir falso positivo e também para a detecção de probe atacks disperos

Future research is directed to the fully automated blind attack detection, extending the evaluation to performance and accuracy analyis of the approach to well-known datasets of network attack detection, and analysing the performance according to parameter configuration in order to better understand the impact of $N, M$ and $Q$ and identify their best configuration for obtaining gains into processing time and acuracy.

>>>>

\section{Complexity Analysis and Performance Evaluation}
\label{sec:Complexity}

This section discusses the computational complexity and the performance evaluation of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.

@@

For better understanding the scalability and impact of configurations of $N$, $M$ and $Q$, we evaluated the processing time requited for different scenarios of parameter configurations and dataset, measuring the processing time of:

\begin{enumerate}
	\item Eigen analysis based on sample covariance of zero mean;
	\item Eigen analysis based on sample covariance of zero mean and unitary standard deviation;
	\item EDC MOS scheme;
\end{enumerate}

The performance evaluation focus on the main steps, which were previously discussed regarding the complexity analysis. The data modeling is also a time consuming step, however its processing can be optimized trough distributed processing techniques, such as MapReduce, achieving high throughput for packet counting or even for deep packet inspection \cite{Vieira2013}.

The experiments were performed in a desktop computer with a Intel Core i7-4510U 2.00GHz and 16 GB of RAM, considering: variations on the network traffic time; the frame size denoted as $N$; the number of network ports denoted as $M$; the mean processing time for eigen analysis based on sample covariance of zero mean, denoted as 1-time; the mean processing time for eigen analysis based on sample covariance of zero mean and unitary standard deviation, denoted as 2-time; and the mean processing time for EDC MOS scheme, denoted as 3-time. The mean time calculations was obtained from 200 measurement repetitions, in order to obtain reliable values.

Table \ref{tab:11} presents the measured results. The experiment considered traffic time of 16, 20 and 22 hours, according to the selected traffic time per day available by the DARPA dataset. It is possible to observe the processing time increases according to the increment in traffic time, around 2 or 3 times for 1-time and 2-time, but the worst measured processing time is 4.7250 milliseconds.

\begin{table*}[!t]
	\caption{Processing time of the main steps for anomaly detection}
	\label{tab:11}
	\centering
	\begin{tabular}{|r|r|r|r|r|r|r|}
		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Traffic Time\\(hour)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Frame Size\\(min)\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Num. Ports\end{tabular}	& \begin{tabular}[x]{@{}l@{}}1-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}2-time\\(ms)\end{tabular}	& \begin{tabular}[x]{@{}l@{}}3-time\\(ms)\end{tabular}	\\ \hline
16	& 10	& 17	& 0.7900	& 0.8100	& 0.0650\\ \hline
16	& 20	& 17	& 0.5250	& 0.5950	& 0.0100\\ \hline
16	& 60	& 17	& 0.9700	& 1.1400	& 0.0250\\ \hline
16	& 120	& 17	& 0.6050	& 0.6100	& 0.0050\\ \hline
16	& 60	& 34	& 1.2750	& 1.2200	& 0.0050\\ \hline
16	& 120	& 34	& 1.1200	& 1.1700	& 0.0050\\ \hline
20	& 10	& 17	& 2.7950	& 2.8950	& 1.1000\\ \hline
20	& 20	& 17	& 2.0700	& 2.0200	& 0.3500\\ \hline
20	& 60	& 17	& 1.0250	& 1.0450	& 0.0650\\ \hline
20	& 120	& 17	& 1.0000	& 1.0700	& 0.0350\\ \hline
20	& 60	& 34	& 2.9650	& 3.2100	& 0.0400\\ \hline
20	& 120	& 34	& 2.9950	& 3.1150	& 0.0200\\ \hline
22	& 10	& 17	& 4.7250	& 4.0850	& 1.4600\\ \hline
22	& 20	& 17	& 2.3200	& 2.6800	& 0.2450\\ \hline
22	& 60	& 17	& 1.0700	& 1.1200	& 0.0300\\ \hline
22	& 120	& 17	& 0.9900	& 1.0500	& 0.0250\\ \hline
22	& 60	& 34	& 3.0850	& 3.1250	& 0.0650\\ \hline
22	& 120	& 34	& 2.8100	& 2.9600	& 0.0250\\ \hline
	\end{tabular}
\end{table*}

It is possible to observe that the processing time increases with the frame size $N$ decreasing, therefore it is possible to evaluate the frame size that produces better identification rates and acceptable processing time. The number of ports evaluated during the proposed scheme is also an important variable regarding processing time optimizations, since the significant increase of processing time observed between scenarios considering 17 or 34 ports, with growth between 7\% and 199\%.

@@

The proposed framework is evaluated and the experimental results show that synflood, fraggle and port scan attacks can be detected accurately and with great detail in an automatic and blind fashion, applying signal processing concepts for traffic modeling and through approaches based on MOS and eigen similarity analysis. The main contributions of this work were: the extension of an approach based on MOS combined with eigen analysis to blindly detect time frames under network attack; the proposal and evaluation of an eigen similarity based framework to identify details of network attacks, presenting accuracy of timely detection and identification of TCP/UDP ports under attack, as well as presenting acceptable complexity and performance regarding the processing time.

@@

Future research is directed to improvements for obtaining better false positive rates, as well as for make the proposed framework able to identify sparse probe attacks. Additionally, future research can evaluate the application of the proposed approach to different domains, considering cases that are aware to behavioral analysis.
